Best practices for forecasting changes in biodiversity: an example using breeding birds

### 2017-08-13 manuscript/draft.Rmd
abstract & title

---
title: Using best practices for forecasting biodiversity changes in breeding birds
author:
- David J. Harris
- Shawn D. Taylor
- Ethan P. White
bibliography: refs.bib
csl: csl.csl
geometry: left=3.5cm, right=3.5cm, top=2.25cm, bottom=2.25cm, headheight=12pt, letterpaper
fontsize: 12pt
header-includes:
    - \usepackage{times}
    - \usepackage{setspace}
    - \doublespacing
    - \usepackage{lineno}
    - \linenumbers
output:
  pdf_document: default
  html_document: default
---

@@

Biodiversity forecasts are important for conservation, management, and
evaluating how well current models characterize natural systems. While there are
an increasing number of studies forecasting biodiversity, there is little
information available on how well these forecasts work. Most diversity forecasts
do not conduct meaningful evaluations of the forecasts, fail to account for
uncertainty, and do not use time-series data that captures the actual dynamics
being studied. Using hindcasting we made and evaluated `r numbers$n_forecasts`
forecasts for breeding bird species richness from each of 6 modeling approaches
at `r numbers$N_predict_sites` sites throughout the continental United States,
following ten suggested best practices for forecasting. Forecasts were made for
up to a decade after the final training observations. While all of the models we
evaluated produce reasonable forecasts, none of them consistently outperforms
the simple baseline of constant long-term average richness. At present, our best
forecast for future richness is thus to assume that it is unchanging, and
current modeling approaches do not reliably improve on this simple baseline.
Most of the best practices implemented in this study directly influence the
outcome of the forecasts and result in different conclusions from those made in
previous studies. This emphasizes the importance of following good practice in
making ecological forecasts and of starting to actively make more forecasts in
such a way that they can be quickly evaluated and improved on.

>>>>

---
title: Best practices for forecasting changes in biodiversity: an example using breeding birds
author:
- David J. Harris
- Shawn D. Taylor
- Ethan P. White
bibliography: refs.bib
csl: csl.csl
geometry: left=3.5cm, right=3.5cm, top=2.25cm, bottom=2.25cm, headheight=12pt, letterpaper
fontsize: 12pt
header-includes:
    - \usepackage{times}
    - \usepackage{setspace}
    - \doublespacing
    - \usepackage{lineno}
    - \linenumbers
output:
  pdf_document: default
  html_document: default
---

@@

Biodiversity forecasts are important for conservation, management, and
evaluating how well current models characterize natural systems. While there are
an increasing number of studies forecasting biodiversity, there is little
information available on how well these forecasts work. Most biodiversity
forecasts are not evaluated in terms of their ability to predict future
diversity, fail to account for uncertainty, and do not use time-series data that
captures the actual dynamics being studied. We compiled a list of ten best
practices for forecasting biodiversity and used hindcasting to evaluate six
different modeling approaches for predicting breeding bird species richness.
Each method was evaluated at `r numbers$N_predict_sites` sites distributed
throughout the continental United States. Forecasts were evaluated for up to a
decade after the final training observations. While all of the models that we
evaluated explained most of the variance in richness, none of them consistently
outperforms a simple baseline model where each site's species richness
fluctuates randomly around a constant value. At present, our best forecast for
future richness is thus to assume that it is unchanging, and current modeling
approaches do not reliably improve on this simple baseline. Most of the best
practices implemented in this study (e.g., evaluating confidence intervals as
well as point estimates) directly influence the outcome of the forecasts; some
even led to large changes in the relative performance of different modeling
approaches. We emphasize the importance of following specific best practices
when making ecological forecasts, and of robust evaluation of model predictions
to facilitate rapid improvement in our biodiversity forecasts.
